---
title: "Case Study_Cyclistic _Bike_Sharing"
author: "Anup  Painuly"
date: "10/3/2021"
output:
  pdf_document: default
  html_document: default
---

# Cyclistic Bike-Sharing Analysis

## Setting up the working environment

```{r}
options(repos = list(CRAN="http://cran.rstudio.com/"))
```


```{r Package Installation, warning=FALSE}
install.packages("tidyverse")
library(tidyverse)
library(lubridate)
library(skimr)
install.packages("dplyr")
library(dplyr)
install.packages("skimr")
install.packages("janitor")
library(janitor)

```

## Importing Data from CSV files

```{r Import Data}
Trips_Apr20 <- read_csv('202004-divvy-tripdata.csv')
spec(Trips_Apr20)
Trips_May20 <- read_csv('202005-divvy-tripdata.csv')
spec(Trips_May20)
Trips_Jun20 <- read_csv('202006-divvy-tripdata.csv')
spec(Trips_Jun20)
Trips_Jul20 <- read_csv('202007-divvy-tripdata.csv')
spec(Trips_Jul20)
Trips_Aug20 <- read_csv('202008-divvy-tripdata.csv')
spec(Trips_Aug20)
Trips_Sep20 <- read_csv('202009-divvy-tripdata.csv')
spec(Trips_Sep20)
Trips_Oct20 <- read_csv('202010-divvy-tripdata.csv')
spec(Trips_Oct20)
Trips_Nov20 <- read_csv('202011-divvy-tripdata.csv')
spec(Trips_Nov20)
Trips_Dec20 <- read_csv('202012-divvy-tripdata.csv')
spec(Trips_Dec20)
Trips_Jan21 <- read_csv('202101-divvy-tripdata.csv')
spec(Trips_Jan21)
Trips_Feb21 <- read_csv('202102-divvy-tripdata.csv')
spec(Trips_Feb21)
Trips_Mar21 <- read_csv('202103-divvy-tripdata.csv')
spec(Trips_Mar21)
Trips_Apr21 <- read_csv('202104-divvy-tripdata.csv')
spec(Trips_Apr21)

```
## Comparing Column names across all the imported files
```{r Coumn name uniformity}

colnames(Trips_Apr20)
colnames(Trips_May20)
colnames(Trips_Jun20)
colnames(Trips_Jul20)
colnames(Trips_Aug20)
colnames(Trips_Sep20)
colnames(Trips_Oct20)
colnames(Trips_Nov20)
colnames(Trips_Dec20)
colnames(Trips_Jan21)
colnames(Trips_Feb21)
colnames(Trips_Mar21)
colnames(Trips_Apr21)

```

## Data frame Inspection for Incongruity

```{r Data frame Inspection}

str(Trips_Apr20)
str(Trips_May20)
str(Trips_Jun20)
str(Trips_Jul20)
str(Trips_Aug20)
str(Trips_Sep20)
str(Trips_Oct20)
str(Trips_Nov20)
str(Trips_Dec20)
str(Trips_Jan21)
str(Trips_Feb21)
str(Trips_Mar21)
str(Trips_Apr21)

```

## Data type Inspection for uniformity

```{r Datatype Uniformity}

compare_df_cols(
  Trips_Apr20,
  Trips_May20,
  Trips_Jun20,
  Trips_Jul20,
  Trips_Aug20,
  Trips_Sep20,
  Trips_Oct20,
  Trips_Nov20,
  Trips_Dec20,
  Trips_Jan21,
  Trips_Feb21,
  Trips_Mar21,
  Trips_Apr21,
  return = "mismatch"
)
4
```
## Converting inconsistent datatype across the tables

```{r}

Trips_Apr20 <- mutate(
  Trips_Apr20,
  end_station_id =
    as.character(end_station_id),
  start_station_id =
    as.character(start_station_id)
)

Trips_May20 <- mutate(
  Trips_May20,
  end_station_id =
    as.character(end_station_id),
  start_station_id =
    as.character(start_station_id)
)

Trips_Jun20 <- mutate(
  Trips_Jun20,
  end_station_id =
    as.character(end_station_id),
  start_station_id =
    as.character(start_station_id)
)

Trips_Jul20 <- mutate(
  Trips_Jul20,
  end_station_id =
    as.character(end_station_id),
  start_station_id =
    as.character(start_station_id)
)

Trips_Aug20 <- mutate(
  Trips_Aug20,
  end_station_id =
    as.character(end_station_id),
  start_station_id =
    as.character(start_station_id)
)

Trips_Sep20 <- mutate(
  Trips_Sep20,
  end_station_id =
    as.character(end_station_id),
  start_station_id =
    as.character(start_station_id)
)

Trips_Oct20 <- mutate(
  Trips_Oct20,
  end_station_id =
    as.character(end_station_id),
  start_station_id =
    as.character(start_station_id)
)

Trips_Nov20 <- mutate(
  Trips_Nov20,
  end_station_id =
    as.character(end_station_id),
  start_station_id =
    as.character(start_station_id)
)

Trips_Dec20 <- mutate(
  Trips_Dec20,
  end_station_id =
    as.character(end_station_id),
  start_station_id =
    as.character(start_station_id)
)

Trips_Jan21 <- mutate(
  Trips_Jan21,
  end_station_id =
    as.character(end_station_id),
  start_station_id =
    as.character(start_station_id)
)


Trips_Feb21 <- mutate(
  Trips_Feb21,
  end_station_id =
    as.character(end_station_id),
  start_station_id =
    as.character(start_station_id)
)

Trips_Mar21 <- mutate(
  Trips_Mar21,
  end_station_id =
    as.character(end_station_id),
  start_station_id =
    as.character(start_station_id)
)


Trips_Apr21 <- mutate(
  Trips_Apr21,
  end_station_id =
    as.character(end_station_id),
  start_station_id =
    as.character(start_station_id)
)

```
## Data type re-inspection for uniformity

```{r Datatype reinspection}

compare_df_cols(
  Trips_Apr20,
  Trips_May20,
  Trips_Jun20,
  Trips_Jul20,
  Trips_Aug20,
  Trips_Sep20,
  Trips_Oct20,
  Trips_Nov20,
  Trips_Dec20,
  Trips_Jan21,
  Trips_Feb21,
  Trips_Mar21,
  Trips_Apr21,
  return = "mismatch"
)
  
```
## Consolidating all the dataframes into one

```{r Consolidation of dataframes}

all_trips <-
  bind_rows(
    Trips_Apr20,
    Trips_May20,
    Trips_Jun20,
    Trips_Jul20,
    Trips_Aug20,
    Trips_Sep20,
    Trips_Oct20,
    Trips_Nov20,
    Trips_Dec20,
    Trips_Jan21,
    Trips_Feb21,
    Trips_Mar21,
    Trips_Apr21
  )

```
## Eliminating non-essential columns
```{r}

all_trips <- all_trips %>% 
  select(-c(start_lat, start_lng, end_lat, end_lng))

```
## Renaming of columns

```{r Column renaming}

all_trips <- all_trips %>% rename(
  trips_id = ride_id,
  ride_type =
    rideable_type
  ,
  start_time = started_at,
  end_time = ended_at
  ,
  from_station_name = start_station_name
  ,
  from_station_id = start_station_id
  ,
  to_station_name = end_station_name
  ,
  to_station_id = end_station_id
  ,
  user_type = member_casual
)

```
## Inspection of consolidated data frame

```{r Inspection of Data Frame}

colnames(all_trips)

dim(all_trips)
head(all_trips)

str(all_trips)

summary(all_trips)
skim(all_trips)

```
## Creating date tables

```{r Date Tables}
all_trips$date <- as.Date(all_trips$start_time)
all_tripsmonth <- format(as.Date(all_trips$date), "%m")
all_trips$day <- format(as.Date(all_trips$date), "%d")
all_trips$year <- format(as.Date(all_trips$date), "%Y")
all_trips$day_of_week <- format(as.Date(all_trips$date), "%A")

```
## Addition of new calculated column for ride duration

```{r adding calculated column}
all_trips$ride_length <-
  difftime(all_trips$end_time, all_trips$start_time)

is.factor(all_trips$ride_length)


all_trips$ride_length <-
  as.numeric(as.character(all_trips$ride_length))
is.numeric

```
## Removing the bad data where bikes were taken out of the dock for quality check by divvy

```{r}

skim(all_trips$ride_length)

all_trips_v2 <- all_trips[!(all_trips$ride_length < 0),]
skim(all_trips_v2)


```
## Conducting descriptive analysis

```{r}

summary(all_trips_v2$ride_length)

```
### Date frame was exported to csv for further analysis and visualisation.

### Refer to the link [Click Here](https://public.tableau.com/app/profile/anup6569/viz/CaseStudy-CyclisticBikeSharing/CyclisticBike-SharingInsightsP1) for insights in Tableau
